### 2016-09-04 mcmc.tex
refs and dois

\paragraph{Ensembles:}
As we discussed (\sectionname~\ref{sec:tuning}), one of the critical challenges in the use of
\MCMC\ is tuning the proposal pdf $q(\pars'\given\pars)$.
In problems in which all of the parameters (all of the components or entries of $\pars$) are somehow ``equivalent'',
or the parameters can be seen as components of a vector in a vector
space, there are \emph{ensemble methods} that make use of not just one
random-walker but instead many walkers to automatically generate a properly tuned
proposal distribution.
These methods make use of the distribution of a set of independent
walkers in the parameter space ($\pars$-space) to gauge the typical step-sizes and
directions at which new proposals should be made, obviating or greatly
reducing the tuning requirements.
We are co-authors on one of these methods,
\project{emcee}\note{\citet{emcee}}, which is popular in the
astrophysical sciences.
It is worty of note that ensemble methods have particular
peculiarities about initialization and convergence diagnostics that
are worth understanding before use.\note{For example, we find
  empirically that it is better to initialize the ensemble in a ball
  that is smaller than the full posterior pdf width, not larger.  For
  another example, the ensemble produces a set of chains that can be
  used to compute the Gelman--Rubin diagnostic \citep{gelmanrubin},
  but this is not a conservative test if the initialization was made
  in a small ball.}
Because ensemble methods are updating independent \MCMC\ chains, they
are often easy to parallelize to make use of multiple cores.

@@

\paragraph{Importance:}
Sometimes it is possible to sample efficiently from a distribution
that is close to the target function $f(\pars)$ you care about but
hard to sample efficiently with \MCMC.
An example is a problem that has a simple (Gaussian, say) likelihood
function times a non-trivial prior:
It is easy to sample from a Gaussian likelihood times a Gaussian
prior, even without \MCMC, but it might be hard to do the same sampling
under a more complex prior.
Importance sampling is a direct sampling method (not really an \MCMC\
method at all):
You take a sampling from the approximate function (which by assumption
was easy) and then you re-weight the sampling (or do a subsequent
rejection of samples) using the ratio of the true function to the
approximate function as a weight or probability.
We use this quite a bit in our hierarchical inferences.\note{%
  For example, we use importance sampling to convert individual-planet
  exoplanet inferences (all performed with dumb priors) into a full-population
  consistent hierarchical model in \citet{eccentricity}.}

@@

\paragraph{Nested:}
Similar to tempering-like methods are a class of methods called ``nested''
samplers, which also smoothly change the target distribution away from $f(\pars)$.
In the nested case, instead of increasing the temperature, the
samplings are of a censored version of the prior, censored by the
value of the likelihoood function.
When the likelihood censoring is strong, only the most high-likelihood
parts of the prior get sampled; when the likelihood censoring is weak,
almost all of the prior gets sampled.
The idea behind nested sampling---like tempering---is that it is
designed to be a good method for exploring the full parameter space or
searching for all of the modes of the posterior pdf.
It is also designed such that it produces not just a sampling but also
an estimate of the integral $Z$ of the likelihood times the prior
(the Bayes factor or fully marginalized likelihood).

@@

@@

  \textit{Astron.\,J.} \textbf{146} 7

@@

  \textit{Comm.\.Appl.\,Math.\,Comp.\,Sci.}, \textbf{5}(1), 65–-80.

>>>>

\paragraph{Ensembles:}
As we discussed (\sectionname~\ref{sec:tuning}), one of the critical challenges in the use of
\MCMC\ is tuning the proposal pdf $q(\pars'\given\pars)$.
In problems in which all of the parameters (all of the components or entries of $\pars$) are somehow ``equivalent'',
or the parameters can be seen as components of a vector in a vector
space, there are \emph{ensemble methods} that make use of not just one
random-walker but instead many walkers to automatically generate a properly tuned
proposal distribution.
These methods make use of the distribution of a set of independent
walkers in the parameter space ($\pars$-space) to gauge the typical step-sizes and
directions at which new proposals should be made, obviating or greatly
reducing the tuning requirements.
We are co-authors on one of these methods,
\project{emcee}\note{\citet{emcee}}, which is popular in the
astrophysical sciences, but there are others\note{One that we are
  interested in is \project{Kombine} (Farr \& Farr, in preparation).}.
It is worty of note that ensemble methods have particular
peculiarities about initialization and convergence diagnostics that
are worth understanding before use.\note{For example, we find
  empirically that it is better to initialize the ensemble in a ball
  that is smaller than the full posterior pdf width, not larger.  For
  another example, the ensemble produces a set of chains that can be
  used to compute the Gelman--Rubin diagnostic \citep{gelmanrubin},
  but this is not a conservative test if the initialization was made
  in a small ball.}
Because ensemble methods are updating independent \MCMC\ chains, they
are often easy to parallelize to make use of multiple cores.

@@

\paragraph{Importance:}
Sometimes it is possible to sample efficiently from a distribution
that is close to the target function $f(\pars)$ you care about, even when the target $f(\pars)$ itself is
hard to sample efficiently with \MCMC.
An example is a problem that has a simple (Gaussian, say) likelihood
function times a non-trivial prior:
It is easy to sample from a Gaussian likelihood times a Gaussian
prior, even without \MCMC, but it might be hard to do the same sampling
under a more complex prior.
Importance sampling is a direct sampling method (not really an \MCMC\
method at all):
You take a sampling from the approximate function (which by assumption
was easy) and then you re-weight the sampling (or do a subsequent
rejection of samples) using the ratio of the true function to the
approximate function as a weight or probability.
We use this quite a bit in our hierarchical inferences.\note{%
  For example, we use importance sampling to convert individual-planet
  exoplanet inferences (all performed with dumb priors) into a full-population
  consistent hierarchical model in \citet{eccentricity}.}

@@

% DFM: Is the last sentence of the above paragraph correct??

\paragraph{Nested:}
Similar to tempering-like methods are a class of methods called ``nested''
samplers\note{For example, see \citet{skilling} and \citet{dnest}},
which also smoothly change the target distribution away from $f(\pars)$.
In the nested case, instead of increasing the temperature, the
samplings are of a censored version of the prior, censored by the
value of the likelihoood function.
When the likelihood censoring is strong, only the most high-likelihood
parts of the prior get sampled; when the likelihood censoring is weak,
almost all of the prior gets sampled.
The idea behind nested sampling---like tempering---is that it is
designed to be a good method for exploring the full parameter space or
searching for all of the modes of the posterior pdf.
It is also designed such that it produces not just a sampling but also
an estimate of the integral $Z$ of the likelihood times the prior
(the Bayes factor or fully marginalized likelihood).

@@

\bibitem[Brewer \etal(2011)]{dnest}
  Brewer,~B.~J., P\'artay,~L.~B. & Cs\'anyi,~G., 2011,
  {Diffusive nested sampling},
  \textit{Stat.\,Comput.}\ \textbf{21} 649
  \doi{10.1007/s11222-010-9198-8}.

@@

  \textit{Astron.\,J.}\ \textbf{146} 7

@@

  \textit{Comm.\.Appl.\,Math.\,Comp.\,Sci.}, \textbf{5}(1), 65–-80
  \doi{10.2140/camcos.2010.5.65}.
